\section{Elaboration of Compilation Steps}

In this section the order and strategy of elaboration of compilation steps will be discussed.

\subsection{Overview of Phases}

\begin{enumerate}
	\item Parsing: per file (text -> AST)
	\item Module AST: merge ASTs into a module AST
	\item Global and Member aggregation: Collect names and types of visible entities defined in global scopes and type/class/... members.
    \item File-Import Resolution: Resolve imports, check for duplicate names in a scope, resolve all-imports. (does not include targets of imports and import expressions)
	\item Name, Type, Import Expression and Operator Resolution: Bind names, infer and check types, turn operator applications into calls. \todo{Das hier ist eine bedarfsgetriebene evaluation von etwas, das in traditionellen Compilern in mehreren pässen erledigt werden würde}
    \item Removal of Unresolved expressions (i.e. check their absence?)
	\item Property checks: Abstract, Native
	\item Access Checks: Check visibility
	\item Removal of Block Parameters: Inline methods taking block parameters.
	\item Removal of Imports: Remove all entities introduced for implicit scope access via imports.
	\item Shadow Checks: Calculate warnings for shadowed entities
	\item IR generation: Create tl-file that can be passed to other compilations or the code generator.
\end{enumerate}

\subsection{Evaluation Order for Name Resolution}

Path imports require that expressions used in the path can be evaluated.
The problem with import paths and operator usage is that new types can be introduced by them and, hence, new dependencies can arise after definining an evaluation order.
\todo{irgendwie habe ich gerade nicht das gefühl, dass das aufgeht; vermutlich muss man die regel aufstellen, dass es keine abhängikeit an das aktuelle modul geben darf, das wäre aber enttäuschend unscharf}


\subsection{File-Import Resolution}

Check for duplicates can be performed without knowing the targets of an import expression, because they end with the name that the target will be refered under.
Names of declared entities are statically known anyway.


\subparagraph{Note}
 This step does not include duplicate names inside of block expressions.


\subsection{Removal of Block Parameters}
This step may introduce illegal usage of internal state.
Hence, access checks must happen before.